---
title: "AnndataR_conversion_protein"
output: html_document
date: "2026-07-29"
---

---
title: "AnndataR"
output: html_document
date: "2026-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(anndataR)
library(renv)
library(Seurat)
```

install BioCManager in renv
```{r}
#if (!requireNamespace("BiocManager", quietly = #TRUE)) {
#  install.packages("BiocManager")
#}
#BiocManager::install("anndataR")
```


```{r}

SeuratObj <- readRDS("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Final_app_objects/Harmony_Treat_Exp_Tech_ADT_Multi_VIS.rds")

```

```{r}
DefaultAssay(SeuratObj) <- "ADT"
SeuratObj  <- JoinLayers(SeuratObj, assay = "ADT")
```


```{r}
adata <- as_AnnData(
  SeuratObj,
  x_mapping     = "counts",
  layers_mapping = c(data= "data"),    # raw counts only -- don't carry over data/scale.data
  obs_mapping   = TRUE,
  obsm_mapping  = FALSE,
  varm_mapping  = FALSE,
  obsp_mapping  = FALSE,
  uns_mapping   = FALSE,
  output_class  = "InMemory"
)
```

Verify that the adata object and the Seurat objects are the same 
```{r}
adata
print("-------------------------")
SeuratObj
```
Seurat stores matrices as features x cells, and AnnData as observations x variables, which explains the transposed dimensioms. The Seurat object holds 10 more features, because it holds another assay: HTO. The RNA layers are the same

```{r}
# Genes in Seurat but not in AnnData
lost_genes <- setdiff(rownames(SeuratObj), rownames(adata$var))
lost_genes

length(lost_genes)

# Genes in AnnData but not in Seurat
extra_genes <- setdiff(rownames(adata$var), rownames(SeuratObj))
extra_genes

sum(duplicated(rownames(SeuratObj)))
sum(duplicated(rownames(adata$var)))
```
These metrics show no differences? So how come the adata object have less dimensions?

```{r}
dim(SeuratObj)
dim(adata)

nrow(SeuratObj)
ncol(SeuratObj)

adata$shape
```

```{r}
# shape
dim(SeuratObj)         # genes × cells
adata$shape      # cells × genes
print("-----------head and tail----------")

# head and tail content
head(colnames(SeuratObj))
head(rownames(adata$obs))

tail(colnames(SeuratObj))
tail(rownames(adata$obs))

print("-----------gene names-----------")
# gene names 
head(rownames(SeuratObj))
head(rownames(adata$var))

tail(rownames(SeuratObj))
tail(rownames(adata$var))
print("-------------expression matrix--------------")


# expression matrix
LayerData(SeuratObj, assay = "RNA", layer = "counts")[1:5, 1:5]

adata$layers[["counts"]][1:5, 1:5]

print("-------------cell metadata--------------")

# cell metadata
head(SeuratObj@meta.data)

head(as.data.frame(adata$obs))

```


```{r}
write_h5ad(
  adata,
  path = "AnnDataR_protein.h5ad",
  mode = "w"
)
```

