---
title: "AnndataR_RNA_and_protein"
output: html_document
date: "2026-07-29"
---

## R Markdown

---
title: "AnndataR"
output: html_document
date: "2026-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Here, the final integrated Seurat object of the Samuels' analysis will be converted to an AnnData object, to later harmonize the dataset using totalVI.

```{r}
library(anndataR)
library(renv)
library(Seurat)
library(Matrix)
library(ggplot2)
```

install BioCManager in renv
```{r}
#if (!requireNamespace("BiocManager", quietly = #TRUE)) {
#  install.packages("BiocManager")
#}
#BiocManager::install("anndataR")
```


```{r}
SeuratObj <- readRDS("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Final_app_objects/Harmony_Treat_Exp_Tech_ADT_Multi_VIS.rds")
```

Proteins present in the object:
```{r}
SeuratObj[["ADT"]]
```

```{r}
rownames(SeuratObj@assays$ADT$data)
```

Filtering is done based on the isotype control I think, so everything below that count is removed

```{r}

assay_name <- "ADT"

print(Layers(SeuratObj[[assay_name]]))   # inspect what's actually there

SeuratObj <- JoinLayers(SeuratObj, assay = assay_name)

print(Layers(SeuratObj[[assay_name]]))   # should now show a single "counts" (+ maybe "data")

counts <- GetAssayData(SeuratObj, assay = assay_name, layer = "counts")
cat("Features:", nrow(counts), "| Cells:", ncol(counts), "\n")

## -----------------------------------------------------------------
## 1. Total counts per feature (summed across all cells), ranked.
## -----------------------------------------------------------------
total_counts <- Matrix::rowSums(counts)
total_counts_sorted <- sort(total_counts)

cat("\n--- Lowest 30 features by total count ---\n")
print(head(total_counts_sorted, 30))

cat("\n--- Summary of total counts across all features ---\n")
print(summary(total_counts))

## -----------------------------------------------------------------
## 2. Flag any isotype/control features still present -- these are
##    your natural background reference. A real marker with a total
##    count at or below an isotype control's total is presumably
##    nonspecific binding, not biology.
## -----------------------------------------------------------------
isotype_like <- grep("igg|isotype|rat ig|hamster ig|streptavidin",
                      rownames(counts), ignore.case = TRUE, value = TRUE)
cat("\n--- Isotype/control-like features detected in this object ---\n")
if (length(isotype_like) > 0) {
  print(total_counts[isotype_like])
} else {
  cat("None found -- you'll need to set a threshold manually / by distribution shape.\n")
}

## -----------------------------------------------------------------
## 3. Plot ranked total counts (log10) to see where signal drops off,
##    with isotype control(s) marked for reference.
## -----------------------------------------------------------------
df <- data.frame(
  feature = names(total_counts_sorted),
  total   = total_counts_sorted,
  rank    = seq_along(total_counts_sorted)
)
df$is_isotype <- df$feature %in% isotype_like

p <- ggplot(df, aes(x = rank, y = total + 1, color = is_isotype)) +
  geom_point() +
  scale_y_log10() +
  scale_color_manual(values = c("FALSE" = "grey40", "TRUE" = "red")) +
  labs(x = "Feature rank (low to high)", y = "Total counts (log10, +1)",
       color = "Isotype/control",
       title = "Ranked ADT total counts per feature") +
  theme_bw()
print(p)

## -----------------------------------------------------------------
## 4. SET YOUR THRESHOLD HERE.
##    Option A (recommended if isotype controls are present): use the
##    max isotype control total as the cutoff.
##    Option B: pick a number by eye from the plot/printed table above.
## -----------------------------------------------------------------
if (length(isotype_like) > 0) {
  threshold <- max(total_counts[isotype_like])
  cat("\nSuggested threshold (max isotype control total counts):", threshold, "\n")
} else {
  threshold <- NA  # <-- set manually, e.g. threshold <- 500
}

# threshold <- 500   # <-- uncomment/edit to override the suggestion above

stopifnot(!is.na(threshold))

## -----------------------------------------------------------------
## 5. Apply the filter and rebuild the assay.
## -----------------------------------------------------------------
keep_features <- names(total_counts)[total_counts >= threshold]
dropped_features <- setdiff(rownames(counts), keep_features)

cat("\nDropping", length(dropped_features), "low-count features:\n")
print(dropped_features)
cat("\nKeeping", length(keep_features), "features.\n")

new_counts <- counts[keep_features, ]

new_assay <- CreateAssay5Object(counts = new_counts)
SeuratObj[[assay_name]] <- new_assay
DefaultAssay(SeuratObj) <- assay_name

```
Verify the filtering
```{r}
rownames(SeuratObj@assays$ADT)
```

Turn it into an anndata object
```{r}
## -----------------------------------------------------------------
## 1. Convert the RNA assay to AnnData. X = raw counts.
##    obsm_mapping = TRUE here will pick up any existing Seurat
##    reductions (PCA/UMAP) if present -- fine to keep.
## -----------------------------------------------------------------
adata <- as_AnnData(
  SeuratObj,
  assay_name    = "RNA",
  x_mapping     = "counts",
  layers_mapping = FALSE,    # raw counts only -- don't carry over data/scale.data
  obs_mapping   = TRUE,
  obsm_mapping  = TRUE,
  varm_mapping  = TRUE,
  obsp_mapping  = TRUE,
  uns_mapping   = TRUE,
  output_class  = "InMemory"
)

adata

## -----------------------------------------------------------------
## 2. Pull out ADT counts and format for obsm: AnnData obsm arrays
##    must be (n_obs x n_features), i.e. cells x proteins -- the
##    TRANSPOSE of how Seurat stores it (features x cells).
## -----------------------------------------------------------------
adt_counts <- GetAssayData(SeuratObj, assay = "ADT", layer = "counts")

## CRITICAL: obsm rows must align 1:1, in the same order, with
## adata$obs (i.e. with colnames(SeuratObj)). If ADT and RNA ever had
## different cells filtered (e.g. you subset one assay but not the
## other), this alignment will silently break. Verify explicitly:
stopifnot(identical(colnames(adt_counts), colnames(SeuratObj)))

adt_mat <- as.matrix(t(adt_counts))   # now cells x proteins

## -----------------------------------------------------------------
## 3. Attach to obsm. Name follows the scvi-tools/totalVI convention
##    ("protein_expression") -- rename if you're using something else.
## -----------------------------------------------------------------
## HDF5 uses '/' internally as a path/group separator, so protein
## names containing a literal '/' (e.g. "F4/80", "TCR g/d-0121")
## break dataset creation when used as column names. Sanitize before
## writing, and keep the original names recoverable in uns.
original_protein_names <- colnames(adt_mat)
safe_protein_names <- gsub("/", "_", original_protein_names, fixed = TRUE)

if (any(safe_protein_names != original_protein_names)) {
  cat("Sanitized protein names containing '/':\n")
  changed <- which(safe_protein_names != original_protein_names)
  print(data.frame(original = original_protein_names[changed],
                    sanitized = safe_protein_names[changed]))
}

colnames(adt_mat) <- safe_protein_names

adt_df <- as.data.frame(adt_mat)

## CRITICAL: explicitly force the row index to the cell barcodes.
## Relying on as.data.frame() to carry rownames through write_h5ad()
## silently failed -- it defaulted to sequential integers ("1", "2", ...)
## which breaks AnnData's requirement that an obsm data.frame's index
## match adata.obs_names exactly (this is what caused the Python
## ValueError on read).
rownames(adt_df) <- colnames(SeuratObj)

## Validate against AnnData's own obs_names BEFORE writing -- fail
## here in R (easy to debug) rather than after writing to h5ad and
## discovering it in Python.
stopifnot(identical(rownames(adt_df), adata$obs_names))

adata$obsm[["protein_expression"]] <- adt_df

## Re-check post-assignment, since the assignment step itself could
## still silently coerce/drop the index depending on how anndataR's
## obsm setter handles data.frames.
stopifnot(identical(rownames(adata$obsm[["protein_expression"]]), adata$obs_names))

## Keep the ORIGINAL (pre-sanitized) names too, so you can map back
## if needed -- order matches the sanitized columns above 1:1.
adata$uns[["protein_names"]] <- original_protein_names
```
Check the AnnData object
```{r}
head(adata$obsm$protein_expression)
head(adata$X)
```


Write out
```{r}
write_h5ad(adata, "C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/AnndataR/AnndataR_RNA_and_protein.h5ad")
```

```{r}
adata_check <- read_h5ad("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/AnndataR/AnndataR_RNA_and_protein.h5ad")
identical(rownames(adata_check$obsm[["protein_expression"]]), adata_check$obs_names)
```

