---
title: "Harmony_protein"
output: html_document
date: "2026-07-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script, we integrate the cDC1 dataset using Harmony, for the orig.ident, experiment and treatment layers.

Before we can run Harmony, we still need to remove 3 proteins with zero counts in the whole assay, and change a typo in the "experiment" layer. The proteins don't need to be removed now, as we will only integrate based on the RNA information. 

```{r}
library("renv")
library("Seurat")
library("harmony")
library("ggplot2")
```

Load in object
```{r}
SeuratObj <- readRDS("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Final_app_objects/Harmony_Treat_Exp_Tech_ADT_Multi_VIS.rds")
```

Below, we will first fix the typo.
```{r}
SeuratObj$experiment <- ifelse(
  SeuratObj$experiment == "CITEseq_LNP_pIC_LNPS",
  "CITEseq_LNP_pIC_LNPs",
  SeuratObj$experiment
)
```
Now, fix the proteins

```{r}
# Assumes your Seurat object is named 'SeuratObj' and protein assay is 'ADT'
# (If your assay is named 'protein', replace "ADT" with "protein")

# 1. Join layers on the ADT assay so all cells are in one contiguous matrix
SeuratObj[["ADT"]] <- JoinLayers(SeuratObj[["ADT"]])

# 2. Now GetAssayData will work smoothly without errors!
protein_counts <- GetAssayData(SeuratObj, assay = "ADT", layer = "counts") 

# 3. Calculate row sums across all cells for each protein
protein_sums <- Matrix::rowSums(protein_counts)

# 4. Identify non-zero proteins
expressed_proteins <- names(protein_sums[protein_sums > 0])
num_removed <- length(protein_sums) - length(expressed_proteins)

cat(sprintf("Removing %d zero-expression proteins.\n", num_removed))

# 5. Subset the protein assay in place
SeuratObj[["ADT"]] <- subset(SeuratObj[["ADT"]], features = expressed_proteins)
```

Next, we will re-run some preprocessing for good practice, before integration
```{r}
DefaultAssay(SeuratObj) <- "ADT"
SeuratObj <- JoinLayers(SeuratObj)

SeuratObj <- NormalizeData(SeuratObj, normalization.method = "CLR", margin = 2)
all.genes <- rownames(SeuratObj)
SeuratObj <- ScaleData(SeuratObj, features = rownames(SeuratObj[["ADT"]]))
SeuratObj <- RunPCA(SeuratObj, 
                     features = rownames(SeuratObj[["ADT"]]),
                     npcs = 50,
                     ndims.print = 1:5, nfeatures.print = 10,
                     assay = "ADT",
                     reduction.name = "ADT_pca_int",
                     reduction.key = "adtPC_int_")
```

Run Harmony. We integrate based on 4 categories: treatment, which holds WT or test; experiment, which holds the different treatments (We know, confusing); orig.ident, which holds the different samples; tech, which holds A or B, referring to A: all treatments except CITEseq_Toxo, and B: CITEseq_Toxo group. 

```{r}
SeuratObj <- RunHarmony(SeuratObj, 
                         group.by.vars =c("treatment","experiment","orig.ident", "tech"), 
                         plot_convergence =TRUE, 
                         reduction.use ="ADT_pca_int",
                         reduction.save ="ADT_harmony",
                         theta=c(4,2,2,6),
                         sigma=0.2, 
                         lambda=1,
                         verbose=TRUE
)

```


Find clusters to visualize results
```{r}
SeuratObj <- FindNeighbors(SeuratObj,reduction = "ADT_harmony",dims = 1:10)
SeuratObj <- FindClusters(SeuratObj, resolution = 1)
SeuratObj <- RunUMAP(SeuratObj,reduction = "ADT_harmony",dims = 1:10,reduction.name = "ADT_harmony_umap")
```

Create plot
```{r, fig.width=12, fig.height=10, dpi=300}
# Plot UMAP colored by batch/treatment variables
p1 <- DimPlot(SeuratObj, reduction = "ADT_harmony_umap", group.by = "treatment") + 
  ggtitle("Harmony_ADT: treatment")

p2 <- DimPlot(SeuratObj, reduction = "ADT_harmony_umap", group.by = "experiment") + 
  ggtitle("Harmony_ADT: experiment")

p3 <- DimPlot(SeuratObj, reduction = "ADT_harmony_umap", group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Harmony_ADT: Clusters")

p4 <- DimPlot(SeuratObj, reduction = "ADT_harmony_umap", group.by = "celltype_new", label = TRUE) + 
  ggtitle("Harmony_ADT: celltype")

p5 <- DimPlot(SeuratObj, reduction = "ADT_harmony_umap", group.by = "tech", label = TRUE) + 
  ggtitle("Harmony_ADT: tech")
p1
p2
p3
p4
p5
```

Save object
```{r}
saveRDS(SeuratObj,"Harmony_protein_integrated.rds")
```