---
title: "RShiny_app_testing"
output: html_document
date: "2026-08-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries
```{r}
library("renv")
library("Seurat")
library("anndataR")
```

object
```{r}
obj <- read_h5ad("c:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/Annotation_Comparison/final_annotation_2026.h5ad", as = "Seurat")
```

# Seurat object with the final celltype annotation made in 2026 during Lode's internship under Clint's guidance.
# This object holds only the RNA layer, as that's the assay used for the downstream analysis. The object holds several annotations, of which the celltype_final_2026
# is the final one. The object holds several integrated layers (scANVI, scVI and several Harmony integrations)
# Of these integrations, the Harmony one of Samuel's integration was used for downstream processing. The UMAP of this integration is also present in the object.
# Furthermore, the RNA assay holds 3 layers: counts, data and X, which all hold the counts (So they are redundant?), no scaled data is present, as that would bloat
# the object.

# The important metadata in the object, is: orig.ident (holds experiment data), seurat_clusters (leiden clusters, or louvain?), experiment (holds the experiment perforned), treatment (Test or WT), tech (A or B, referring to 3' or 5' sequencing I think, only Toxo has a different sequencing method), leiden_1 and leiden_2 (referring to leiden resolution used), and celltype_final_2026 

# Additional information calculated in the downstream analysis were: 
# - Conserved markers (significant across all celltypes) (C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/Downstream_processing/conserved_markers_summary.pdf)

# - Marker list, just all genes with a LogFC, pval and how many times significant (C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/Downstream_processing/all_ranked_markers_summary.xlsx)

# - TF scores using the pseudobulk samples (C:\Users\irc\Desktop\Internship Bioinformatics 2025-2026\Lode\Downstream_processing\TF)
# - GSEA and ORA analysis for each celltype in each experiment (C:\Users\irc\Desktop\Internship Bioinformatics 2025-2026\Lode\Downstream_processing\GSEA and C:\Users\irc\Desktop\Internship Bioinformatics 2025-2026\Lode\Downstream_processing\ORA)

# - DESeq2 results from LNP vs WT, Toxo vs LNP and Toxo Test vs WT (C:\Users\irc\Desktop\Internship Bioinformatics 2025-2026\Lode\Downstream_processing\DESeq2)

```{r}
obj@meta.data
```

The object did not fix the SAM06 mistake
```{r}
obj$treatment[obj$orig.ident == "SAM06"] <- "WT"

table(obj$orig.ident[obj$orig.ident == "SAM06"],
      obj$treatment[obj$orig.ident == "SAM06"])
```

Write out
```{r}
saveRDS(obj, "RShiny_tool_object.rds")
```

