---
title: "UCell test"
output: html_document
date: "2026-08-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("renv")
library("Seurat")
library("UCell")
```

Load in data
```{r}
SeuratObj <- readRDS("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/Internship-Lode-Usja/RShiny/App/RShiny_tool_object.rds")
```

gene sets and scores. This block takes too long, so we will inspect the Seuratobject, to see why it takes to long
```{r}
# gene.sets <- list(Late_Mature_signature = c("Xcr1","Cxcl9","Cxcl10","Fscn1"))
# scores <- ScoreSignatures_UCell(GetAssayData(SeuratObj, layer = "data"), features=gene.sets)
# head(scores)
```

Check for duplicates, NAs... in the 3 layers
```{r}
# all have the same dimensions
counts_layer <- GetAssayData(SeuratObj, assay = "RNA", layer = "counts") # raw counts
data_layer   <- GetAssayData(SeuratObj, assay = "RNA", layer = "data") # log normalized
X_layer      <- GetAssayData(SeuratObj, assay = "RNA", layer = "X") # 
```


```{r}
# Class and dimensions
message("counts: ", paste(class(counts_layer), collapse=","), " | dim: ", paste(dim(counts_layer), collapse=" x "))
message("data:   ", paste(class(data_layer), collapse=","),   " | dim: ", paste(dim(data_layer), collapse=" x "))
message("X:      ", paste(class(X_layer), collapse=","),      " | dim: ", paste(dim(X_layer), collapse=" x "))

# Are counts and data identical? (they shouldn't be if data is log-normalized)
identical(counts_layer, data_layer)

# Value range comparison -- normalized data should have small decimals,
# raw counts should be whole numbers (often larger integers)
message("counts range: ", paste(range(counts_layer@x), collapse=" - "))
message("data range:   ", paste(range(data_layer@x), collapse=" - "))
message("X range:      ", paste(range(X_layer@x), collapse=" - "))

# Are all values in "data" whole numbers? (TRUE would mean it's NOT normalized)
message("data layer looks like raw counts (all whole numbers)? ",
        all(data_layer@x == round(data_layer@x)))

# Direct value-by-value comparison on a small subset, for a sanity check
identical(counts_layer[1:20, 1:20], data_layer[1:20, 1:20])
identical(data_layer[1:20, 1:20], X_layer[1:20, 1:20])
```
```{r}
# 1. Duplicate gene names -- never actually checked yet, still a real candidate
message("Duplicate gene names: ", sum(duplicated(rownames(counts_layer))))

# 2. Duplicate cell barcodes
message("Duplicate cell barcodes: ", sum(duplicated(colnames(counts_layer))))

# 3. Layer join status -- also never confirmed
print(Layers(SeuratObj, assay = "RNA"))

# 4. NA/Inf check on the data layer specifically
message("Any NA in data@x: ", anyNA(data_layer@x))
message("Any Inf in data@x: ", any(is.infinite(data_layer@x)))
```

```{r}
# ucell_ranks_cache <- "C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/Internship-Lode-Usja/RShiny/RShiny/App/data/ucell_ranks.rds"
# ucell_ranks <- readRDS(ucell_ranks_cache)

data_test <- data_layer[, sample(ncol(data_layer), 50000)]
signatures <- list(Late_Mature = c("Xcr1", "Cxcl9", "Cxcl10", "Fscn1"))
#    using the SAME ucell_ranks your app already built, run in a totally clean script
#    (no Shiny, no BPPARAM) -- isolates whether the slowness is really about the data,
#    or about something in the app/reactive/parallel context
t0 <- Sys.time()
test_scores <- ScoreSignatures_UCell(data_test,
  features = signatures,
  maxRank = 1500,
  BPPARAM = BiocParallel::SnowParam())

message("Serial scoring (cached ranks, clean script): ",
        round(difftime(Sys.time(), t0, units = "secs"), 1), " sec") 
```
investigate the ranks
```{r}
ucell_ranks
```
Conclusion: all seems fine. The data is a sparse matrix, the ranked file contains the right kind of data (ranks), packages work and the max_rank is the same as pyUCell (default to 1500). So in conclusion, I have no idea why this is so slow in R (approx 15-20 minutes for scoring 4 genes using pre-calculated ranks), as it only takes 3-5 minutes in Python, not using pre-calculated ranks. 
Session info
```{r}
sessionInfo()
```

