---
title: "Harmony_Samuel_anndataR"
output: html_document
date: "2026-08-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Here, the final integrated Seurat object of the Samuels' analysis will be converted to an AnnData object, to later harmonize the dataset using scVI.

```{r}
library(anndataR)
library(renv)
library(Seurat)
```

Load in object
```{r}
SeuratObj <- readRDS("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Final_app_objects/SeuratObjT_Harmomy_Treat_Exp_V2.rds")
```

Make the working object (for AnndataR) smaller
```{r}
SeuratObj_small <- DietSeurat(
    SeuratObj,
    assays = "RNA",
    layers = c("counts", "data"),
    dimreducs = c("harmony", "harmony_umap", "RNA_pca_int", "RNA_umap"),
    graphs = "neighbors"
)
```

```{r}
saveRDS(SeuratObj_small, "data/Samuel_data_Harmony_small.rds")
```

Load in smaller file
```{r}
SeuratObj_small <- readRDS("C:/Users/irc/Desktop/Internship Bioinformatics 2025-2026/Lode/Harmony/Harmony_AnndataR_conversion/data/Samuel_data_Harmony_small.rds")
```

```{r}
DefaultAssay(SeuratObj_small) <- "RNA"
```

```{r}
SeuratObj_small <- JoinLayers(SeuratObj_small, assay = "RNA")
```

Conversion to adata
```{r}
adata <- as_AnnData(
  SeuratObj_small,
  assay_name  = "RNA",
  x_mapping   = "counts",     # or "counts", whichever you use downstream
  obsm_mapping = list(
    RNA_pca_int      = "RNA_pca_int",       # LHS = obsm key you want, RHS = Seurat reduction name
    RNA_umap         = "RNA_umap",
    RNA_harmony      = "harmony",
    RNA_harmony_umap = "harmony_umap"
  )
)
```


```{r}
adata
```

```{r}
head(adata$obsm)
head(str(adata$obsm[["RNA_harmony"]]))
head(dim(adata$obsm[["RNA_harmony"]]) )  # should be n_cells x n_dims, not NULL
```

Verify that the adata object and the Seurat objects are the same 
```{r}
adata
print("-------------------------")
SeuratObj
```
Seurat stores matrices as features x cells, and AnnData as observations x variables, which explains the transposed dimensions. The Seurat object holds 10 more features, because it holds another assay: HTO. The RNA layers are the same

```{r}
# Genes in Seurat but not in AnnData
lost_genes <- setdiff(rownames(SeuratObj), rownames(adata$var))
lost_genes

length(lost_genes)

# Genes in AnnData but not in Seurat
extra_genes <- setdiff(rownames(adata$var), rownames(SeuratObj))
extra_genes

sum(duplicated(rownames(SeuratObj)))
sum(duplicated(rownames(adata$var)))
```
These metrics show no differences? So how come the adata object have less dimensions?

```{r}
dim(SeuratObj)
dim(adata)

nrow(SeuratObj)
ncol(SeuratObj)

adata$shape
```

```{r}
# shape
dim(SeuratObj)         # genes × cells
adata$shape      # cells × genes
print("-----------head and tail----------")

# head and tail content
head(colnames(SeuratObj))
head(rownames(adata$obs))

tail(colnames(SeuratObj))
tail(rownames(adata$obs))

print("-----------gene names-----------")
# gene names 
head(rownames(SeuratObj))
head(rownames(adata$var))

tail(rownames(SeuratObj))
tail(rownames(adata$var))
print("-------------expression matrix--------------")


# expression matrix
LayerData(SeuratObj, assay = "RNA", layer = "counts")[1:5, 1:5]

adata$layers[["counts"]][1:5, 1:5]

print("-------------cell metadata--------------")

# cell metadata
head(SeuratObj@meta.data)

head(as.data.frame(adata$obs))

```


```{r}
write_h5ad(
  adata,
  path = "data/Harmony_RNA_Samuel_anndataR.h5ad",
  mode = "w"
)
```



